---
title: "PanelMatch"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
how to do bibliography
https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Abstract

# Introduction


# DisplayTreatment


## Data Requirements and Assumptions 

```{r, fig.width=5, fig.height=5}
library(PanelMatch)
DisplayTreatment(unit.id = "wbcode2",
                 time.id = "year", legend.position = "none",
                 xlab = "year", ylab = "Country Code",
                 treatment = "dem", data = dem)
```

## Customizing DisplayTreatment Plots

# PanelMatch

## Data Requirements and Assumptions 

## Choosing parameters
```{r}
PM.results <- PanelMatch(lag = 4, time.id = "year", unit.id = "wbcode2", 
                         treatment = "dem", refinement.method = "mahalanobis", 
                         data = dem, match.missing = T, 
                         covs.formula = ~ I(lag(tradewb, 1:4)) + I(lag(y, 1:4)), 
                         size.match = 5, qoi = "att" ,outcome.var = "y",
                         lead = 0:4, forbid.treatment.reversal = FALSE)
```
## Understanding PanelMatch objects

## Understanding matched.set objects

### Subsetting and manipulation of matched.set objects

### Summary function for matched.set objects

### Plot function for matched.set objects

## Checking Covariate Balance

### Using get_covariate_balance

### Using balance_scatter

## Checking Multiple Configurations

# PanelEstimate

## Choosing Parameters

```{r}
PE.results <- PanelEstimate(sets = PM.results, data = dem)
```
## Understanding PanelEstimate Results


### Summary 

### Plot


# Concluding Remarks




